---
title: "Model of acoustic loss"
author: "W. Bauer"
date: "03/31/2015"
output: html_document
---

Though based on optimistic assumptions, a value of 1 kPa was [obtained](energy_dnsity.html) for acoustic pressure (a.k.a. energy density) generated by relaxation of dyes excited by externally applied NIR radiation. This is a reasonably good value, hence acoustic transmission losses between stained site and scalp are of interest.

Here we develop a back-of-the-envelope acoustic transmission model applicable to the BrainWeb discrete phantom. The general technique is "voxelized ray traversal," by which we mean that a "ray" is represented by the sequence of voxels which intersect a given line. As ultrasound traverses the ray, it accumulates losses due to absorption in voxel interiors and reflection at voxel boundaries. Reflected energy is ignored, even though it could reverberate to an extent. Angles of incidence and refraction are also ignored, though they would have significant effects at boundaries between bone and soft tissue. Refinements would be possible if called for.

In addition to losses along a ray, there will be inverse square losses due to spherical radiation from the stained site. We will account for these later.

Acoustic properties for the phantom's 11 tissue types were guesstimated [here](acoustic_properties.html) and recorded in csv form [here](../data/acoustic_properties_bw.csv). A range was given for each property. To begin, random values from each range will be chosen for each tissue type (and for air.)

```{r}
properties <- read.table("../data/acoustic_properties_bw.csv", sep=",", header=TRUE)
rho <- runif(12, properties[,"rho.low"], properties[,"rho.high"])
speed <- runif(12, properties[,"c.low"], properties[,"c.high"])
alpha <- runif(12, properties[,"alpha.low"], properties[,"alpha.high"])
```

Absorption coefficients, alpha, are given in units of $dB/m/MHz$. Since the phantom's voxels are 1 mm on a side, we'll convert to $dB/mm/MHz$. 

```{r}
alpha <- alpha/1000
```

The other necessary quantities are transmission coefficients, also expressed in decibels of loss. These are most easily arranged in a matrix. The general formula for transmitted to (normally incident) power is$$10 log_{10}\left(\frac{4Z_1Z_2}{(Z_1 + Z_2)^2}\right),$$where $Z_i = \rho_i c_i.$ There is no reflection between tissues with identical properties, hence some matrix elements will be zero.

```{r}
Z <- rho*speed
rc <- matrix(0, 12, 12)
for(i in 1:12){
  for(j in 1:12){
    if(Z[i] != Z[j]){
      rc[i,j] <- 10*log10(4*Z[i]*Z[j]/(Z[i]+Z[j])^2)
    }
  }
}
```

We'll use a small section of a sagittal slice for testing.

```{r}
fname <- "../data/subject04_crisp_v.rawb"
phantom <- readBin(fname, what="raw", n=362*434*362, size=1, signed=FALSE, endian="big")
dim(phantom) <- c(362, 434, 362)
slice <- phantom[100,,]
slice <- as.integer(slice)
dim(slice) <- c(434, 362)
rm(phantom)
section <- slice[280:320, 250:310]
section[10,11] # gray
section[30, 39] # muscle/skin next to scalp
temp <- findVoxelCrossings(c(0,10-.5,11-.5), c(0,30-.5,39-.5))
ray <- matrix(c(0,10,11), 1, 3)
for(n in seq(1, nrow(temp))){
  nxt <- ray[nrow(ray),]
  k <- temp[n,4]
  nxt[k] <- 1 + nxt[k]
  ray <- rbind(ray, nxt)
}
points(ray[,2], ray[,3], pch=15, col="white")
```

