---
title: "Discrete Methods"
author: "W. Bauer"
date: "12/18/2014"
output: html_document
---

In the previous note, [Spatial Resolution](spatial_resolution.Rmd), a very simplified model of NIR transmission from cerebral cortex to scalp was shown to be a standard linear system, amounting to a 2D spatial low-pass filter with very small eigenvalues at spatial frequencies above 1 cycle/cm. This would imply extreme difficulty in achieving spatial resolution of better than 1 cm.

Of course, this simplified model's main virtue is its simplicity. Its accuracy is very much open to question, but it may highlight issues to check with more realistic models and data as these become available.

In the previous note, accuracy of numerical integration was a limiting factor. This note uses a discrete form of the model, replacing integrals with sums, both in an attempt to avoid the integration issue and to facilitate quick numerical experiment.

[Nyquist's Sampling Theorem](http://en.wikipedia.org/wiki/Nyquist–Shannon_sampling_theorem) implies that spatial resolution of 3 mm requires an array of detectors spaced at intervals no greater than 1.5 mm on a grid covering the entire area of scalp at which significant signal strength is expected. In what follows, a grid of 1 mm intervals is used.

```{r echo=FALSE, fig.show='hold', fig.align='center'}
# Estimates fraction of photons which will reach the scalp at a point r mm distant
# from the point directly over their origin, assuming their origin is b mm beneath
# the scalp.
f_b <- function(r, b){
  r1 <- sqrt(r^2+b^2)
  b*.92^r1/(4*pi*r1^1.5)
}
integrand <- function(r)r*f_b(r, 10)
```

A discrete model amenable to treatment with a fast Fourier transform, must be finite in spatial extent. Numerical integration shows that the 2D spatial filter derived previously has little relative weight beyond 128 mm from its center:$$\frac{\int_{128}^{\infty}.92^{\sqrt{r^2 + b^2}}\frac{b}{4 \pi (r^2 + b^2)^{1.5}}rdr}{\int_0^{\infty}.92^{\sqrt{r^2 + b^2}}\frac{b}{4 \pi (r^2 + b^2)^{1.5}}rdr} = 1.88 \times 10^{-5}$$A 256x256 grid will cover this radius. However, a fast Fourier transform operates on a periodic or circular grid. To avoid aliasing from wrap-around, linear dimensions should be doubled. Thus a 512x512 grid is used here.

```{r echo=FALSE, fig.show='hold', fig.align='center'}
filter <- matrix(0,512,512)
for(i in 0:128){
  for(j in 0:128){
    r <- sqrt(i^2 + j^2)
    if(r <= 128){
      temp <- f_b(r, 10)
      filter[256-i, 256-j] <- temp
      filter[256+i, 256-j] <- temp
      filter[256-i, 256+j] <- temp
      filter[256+i, 256+j] <- temp
    }
  }
}
filter_spectrum <- fft(filter)/512
eigenval_3mm_res <- abs(filter_spectrum[512, 170])
eigenval_10mm_res <- abs(filter_spectrum[512, 51])
eigenval_max <- max(abs(filter_spectrum))
err_on_inverse <- max(abs(filter - fft(filter_spectrum, inverse=TRUE)/512))
```

Since the spatial grid has 512 1 mm intervals in each dimension, the grid of a corresponding Fourier transform will be in units of 1/512 cycles/mm in each dimension. The spatial resolution of interest correponds to 1/3 cycles/mm, hence to a point at position 512/3 = 171 on either dimension of the Fourier transform grid. The absolute value of the eigenvalue at such a point is 9.267515e-11. This is about 8 orders of magnitude below the maximum of 0.01178317, and is not likely to be entirely inaccurate, since the inverse transform matches the original to within 16 decimal places. A logarithmic plot of eigenvalue magnitudes vs spatial frequencies follows.

```{r echo=FALSE, fig.show='hold', fig.align='center'}
freqs <- (1:256)/512
plot(freqs, abs(filter_spectrum[512, 1:256]), type='l', lwd=2, log='y', xlab="spatial frequency in cycles/mm", ylab="absolute value of eigenvalue", main="Eigenvalue Spectrum of Simple Model")
linear_region <- (floor(.05*512)):(floor(.2*512))
evs <- log(abs(filter_spectrum[512, linear_region]))
frqs <- freqs[linear_region]
fit <- lm(evs ~ frqs)
lines(frqs-.01, exp(fit$fitted.values), lwd=2, lty=2, col="red")
factor_of_10 <- log(10)/coef(fit)[2]
text(.2, 1e-05, "Factor of 10 decrease\nper 1/30 cycles/mm", col="red")
```

In the linear region, eigenvalue magnitudes decrease by a factor of 10 every 1/30 cycle/mm.

A representative eigenvalue correponding to 10 mm resolution has absolute value 1.657966e-05 which is less than 3 orders of magnitude below the maximum, suggesting as in the previous note a limiting spatial resolution of 1 cm.



[download Rmd which includes source code for computations](discrete_methods.Rmd)
