---
title: "Reconstruction by Linear Regression"
author: "W. Bauer"
date: "12/26/2014"
output: html_document
---

### Summary

For a very simplified model of NIR emission and transport, recovery of detail is limited by the size of eigenvalues at spatial frequencies corresponding to the reciprocal of detail size. Simulations suggest that, noise-to-signal (amplitude) ratios will generally obscure spatial frequencies whose eigenvalues are one or two orders of magnitude smaller than the noise-to-signal ratio itself. Since eigenvalues decrease rapidly with spatial resolution, this would pose a fairly stringent requirement for measurement and detection.

### Reconstruction by linear regression

In [Discrete estimation of eigenvalues and spatial resolution](discrete_methods.html) it was shown that, for a very simplified model of NIR emission and transport, recovery of a pattern of depolarization on the cerebral cortex would be very sensitive to noise added to a pattern measured at the scalp. Moreover, it was shown that sensitivity to noise would worsen rapidly with desired spatial resolution, roughly by a factor of 1000 per 1/10 mm of resolution over about half the spatial frequency spectrum. The relevant figure is reproduced below.

```{r echo=FALSE}
# Estimates fraction of photons which will reach the scalp at a point r mm distant
# from the point directly over their origin, assuming their origin is b mm beneath
# the scalp.
f_b <- function(r, b) {
  r1 <- sqrt(r^2+b^2)
  b*.92^r1/(4*pi*r1^1.5)
  }

# Derives the cerebral cortex to scalp transfer function as a 512x512 fft.
make_fwd_xfer_fct <- function(){
  filter <- matrix(0,512,512)
  for(i in 0:128){
    for(j in 0:128){
      r <- sqrt(i^2 + j^2)
      if(r <= 128){
        temp <- f_b(r, 10)
        filter[256-i, 256-j] <- temp
        filter[256+i, 256-j] <- temp
        filter[256-i, 256+j] <- temp
        filter[256+i, 256+j] <- temp
        }
      }
    }
  return(fft(filter)/512)
}

fwd_xfer_fct <- make_fwd_xfer_fct()

# Computes relative magnitude of eigenvalue corresponding to given resolution
eigenv <- function(resolution){
  return(Mod(fwd_xfer_fct[1, floor(512/resolution)]))
}

# Moves the origin from the corners of a 2D array to its center
center_image <- function(img){
  d <- dim(as.matrix(img))
  ctr <- d/2
  # swap blocks on main diagonal
  temp <- img[(1+ctr[1]):d[1], (1+ctr[2]):d[2]]
  img[(1+ctr[1]):d[1], (1+ctr[2]):d[2]] <- img[1:ctr[1], 1:ctr[2]]
  img[1:ctr[1], 1:ctr[2]] <- temp
  # swap blocks on off-diagonal
  temp <- img[1:ctr[1], (1+ctr[2]):d[2]]
  img[1:ctr[1], (1+ctr[2]):d[2]] <- img[(1+ctr[1]):d[1], 1:ctr[2]]
  img[(1+ctr[1]):d[1], 1:ctr[2]] <- temp
  return(img)
}

# Creates a checkerboard pattern with given cell size (in mm.)
# Polarized regions have relative emission rates of 1.32,
# depolarized regions relative rates of 1, and remaining
# areas have rates of 0.
checkerboard <- function(diameter, cell_size){
  radius <- diameter/2
  board <- matrix(0, 512, 512)
  for(i in 1:diameter){
    for(j in 1:diameter){
      if(sqrt((i-25.5)^2 + (j-25.5)^2) <= radius){
        polarized <- (floor(i/cell_size) + floor(j/cell_size))%%2
        if (polarized){
          board[i-25+256,j-25+256] <- 1.32
          } else {
            board[i-25+256,j-25+256] <- 1.0
            }
        }
      }
    }
  return(board)
  }

# Creates expected value of fft of an emitting region of given diameter,
# assuming adjacent cells are independently polarized or depolarized.
checkerboard_mean <- function(diameter){
  radius <- diameter/2
  board <- matrix(0, 512, 512)
  for(i in 1:radius){
    for(j in 1:radius){
      if(sqrt((i-25.5)^2 + (j-25.5)^2) <= radius){
        board[i,j] <- 1.16
        }
      }
    }
  return(fft(board)/512)
}

# Computes the power spectrum (i.e., variance) of the deviation of
# an emitting region from its mean, assuming adjacent emitting cells 
# are independently polarized or depolarized.
# By the independence assumption, the power spectrum is a constant.
checkerboard_power <- function(diameter){
  pwr <- 0
  radius <- diameter/2
  for(i in 1:radius){
    for(j in 1:radius){
      if(sqrt((i-25.5)^2 + (j-25.5)^2) <= radius){
        pwr <- pwr + .16^2
        }
      }
    }
  return(pwr/512/512)
}

# Give the scalp image of a checkerboard of given diameter, and a noise_to_signal
# (amplitude) ratio, reconstruct the checkerboard by linear regression, which
# is optimal under independence assumptions.
recover_checkerboard <- function(board, diameter, noise_to_signal){
  cb_power <- checkerboard_power(diameter)
  # form fft of scalp pattern
  scalp_fft <- fwd_xfer_fct*fft(board)/512
  # power ratio is the square of amplitude ratio
  noise_power <- noise_to_signal^2*mean(Mod(scalp_fft)^2)
  scalp_fft <- scalp_fft + matrix(rnorm(512*512, 0, sqrt(noise_power)), 512, 512)
  # form denominator of optimal inverse filter
  dinv <- fwd_xfer_fct + noise_power/cb_power/Conj(fwd_xfer_fct)
  # use it to estimate fft of original pattern
  cc_fft <- scalp_fft/dinv + (1 - fwd_xfer_fct/dinv)*checkerboard_mean(diameter)
  rm(dinv, scalp_fft)
  # take the inverse transform
  return(Mod(fft(cc_fft, inverse=TRUE)/512))
}

display_centered_image <- function(img, diameter=50, main=""){
  ctr<- dim(img)/2
  radius <- diameter/2
  rws <- seq(ctr[1]-radius, ctr[1]+radius)
  cls <- seq(ctr[2]-radius, ctr[2]+radius)
  image(rws-ctr[1], cls-ctr[1], img[rws, cls], col=terrain.colors(12), main=main, xlab="mm", ylab="mm")
}

check <- function(img, diameter=100){
  display_centered_image(Mod(center_image(img)), diameter)
}

```

```{r echo=FALSE, fig.show='hold', fig.align='center'}
freqs <- (1:256)/512
plot(freqs, Mod(fwd_xfer_fct[512, 1:256]), type='l', lwd=2, log='y', xlab="spatial frequency in cycles/mm", ylab="absolute value of eigenvalue", main="Eigenvalue Spectrum")
linear_region <- (floor(.05*512)):(floor(.2*512))
evs <- log(abs(fwd_xfer_fct[512, linear_region]))
frqs <- freqs[linear_region]
fit <- lm(evs ~ frqs)
lines(frqs-.01, exp(fit$fitted.values), lwd=2, lty=2, col="red")
factor_of_10 <- log(10)/coef(fit)[2]
text(.21, 1e-05, "A factor of 1000 decrease\nper 1/10 cycles/mm", col="red")
```

In that note no attempt at optimal recovery was made, but recovery by linear regression was mentioned. Assuming patterns of depolarization are spatially uncorrelated, (a pessimistic assumption,) recovery by linear regression is straightforward. For this method, simulations pictured below indicate that recovery of details at $n$ mm resolution generally require noise-to-signal amplitude ratios at the scalp no more than 1 or 2 orders of magnitude larger than that of the eigenvalue at spatial frequency $\frac{1}{n}$ cycles/mm.

```{r echo=FALSE, fig.align='center', fig.show='hold'}
plot_recovery <- function(diameter, cell_size, resolution=cell_size, noise_to_signal=eigenv(resolution)*100){
  par(mfrow=c(1,2))
  board <- checkerboard(diameter, cell_size)
  recovery <- recover_checkerboard(board, diameter, noise_to_signal)
  display_centered_image(board, diameter=diameter, main= paste("Checkerboard depolarization\npattern with cell size ", cell_size, "\n(eigenvalue ", signif(eigenv(resolution), 4),")"))
  display_centered_image(recovery, diameter=diameter, main=paste("Recovered pattern,\nnoise-to-signal\namplitude ratio", signif(noise_to_signal, 4)))
  par(mfrow=c(1,1))
}
```

First consider a checkerboard pattern 50 mm in diameter with cell 10 mm on a side. Cells are alternately polarized (white,) and depolarized (red.) There are no emissions from the area colored green. (See the [previous note](discrete_methods.html) for details.) With a noise-to-signal amplitude ratio one order of magnitude greater than the eigenvalue at spatial frequency 1/10 cycles/mm, the checkerboard pattern is discernible.

```{r echo=FALSE, fig.align='center', fig.show='hold'}
plot_recovery(50, 10, noise_to_signal = eigenv(10)*10)
```

In this case, if the noise-to-signal ratio is two orders of magnitude above the appropriate eigenvalue, the checkerboard is not discernible.

```{r echo=FALSE, fig.align='center', fig.show='hold'}
plot_recovery(50, 10, noise_to_signal=eigenv(10)*100)
```

At 4 mm resolution, two orders of magnitude can be tolerated. However, the eigenvalue itself is extremely small. The noise-to-signal ratio's reciprocal will almost surely exceed the number of photons per square mm reaching the scalp.  

```{r echo=FALSE, fig.align='center', fig.show='hold'}
plot_recovery(50, 4, noise_to_signal=eigenv(4)*100)
```

At 3 mm resolution the situation is similar.

```{r echo=FALSE, fig.align='center', fig.show='hold'}
plot_recovery(50, 3, noise_to_signal=eigenv(3)*1e2)
```

[download Rmd which includes source code for computations](optimal_reconstruction.Rmd)