---
title: "Simulated VSD spectra"
author: "W. Bauer"
date: "May 31, 2015"
output: html_document
---

This note is to specify an approximate model of VSD absorption and re-emission and the variation of those properties with membrane potential. Until exact data becomes available, I'll rely on a stylized form  the example shown below.

<br/><br/><center>![spectrum][loew]</center>
Source: L. M. Loew [Design and Use of Organic Voltage Sensitive Dyes](http://www.newbooks-services.de/MediaFiles/Texts/8/9781441965578_Excerpt_001.pdf) (ungated pdf).<br/><br/>

There is a slight skew in the above curves, as would be expected if the frequency spectrum, i.e., relative fluorescence as a function of frequency rather than as a function of wavelength, were unimodal and symmetric. The shapes in the figure can, in fact, be approximated by using a Gaussian frequency spectrum with a mean corresponding to an (arbitrary) peak wavelength of 675 nm, i.e., a frequency of $\frac{300}{675}\:PHz$, and supposing that 150 nm corresponds to 2.5 standard deviations in frequency, i.e.,$$sd = \frac{1}{2.5}\left(\frac{300}{675-125}-\frac{300}{675}\right).$$

```{r fig.align='center', fig.show='hold', fig.width=7.5, echo=FALSE}
fmean <- 300/675
fsd <- (300/525 - fmean)/2.5
# Changing the variable from frequency to wavelength introduces the factor of 300/lambda^2
spec <- function(lambda, shift=0)dnorm(300/(lambda-shift), fmean, fsd)*300/(lambda-shift)^2
lambda <- seq(550, 850, length.out = 100)
plot(lambda, spec(lambda+20), type='l', col="gray", lwd=3, xlab="wavelength (nm)", ylab="", yaxt="n", bty='n', main = "Stylized VSD spectra")
lines(lambda, spec(lambda), col="black", lwd=3)
legend('topright', c("Polarized", "Depolarized"), lwd=3, col=c("black", "gray"), bty='n')
```

Peak emission and absorption wavelengths for various lipid-bound chromophores are given in Leslie Loew's review, (*ibid*.) Spectrum bandwidths are not indicated in the review but widths of several hundred nanometers seem common elsewhere in the literature. Hence, my choice of +/- 150 nm above.

A convenient model based on the above can be created with the following code. Its use and interpretation are discussed below. 

```{r}
# peak -- peak wavelength in nanometers
# width -- width (support) of the spectrum in nanometers
# Returns a function which gives relative fluorescence or absorbance as a function of
# wavelength and shift due to depolarization.
vsdSpectrum <- function(peak, width=300){
   # Create a function for a curve with a 675 nm peak
  # NOTE: the factor 300/lambda^2 comes from a change of variable
  # from frequency to wavelength.
  fsd <- (300/(675-.5*width) - 300/675)/2.5
  fcal <- function(lambda)dnorm(300/lambda, 300/675, fsd)*300/lambda^2
  # Return a function which translates the calibrated curve
  # to the correct peak wavelength and shift, and normalizes
  # it so its peak value is 1.
  function(lambda, depolarization_shift=0){
    fcal(lambda + depolarization_shift + 675 - peak)/fcal(675)
  }
}
```


[loew]: ../images/loew.png