---
title: "Simulated VSD spectra"
author: "W. Bauer"
date: "May 31, 2015"
output: html_document
---

### Stylized form of VSD spectra

This note is to specify an approximate model of VSD absorption and re-emission and the variation of those properties with membrane potential. Until exact data becomes available, I'll rely on a stylized form as shown the example below.

<br/><br/><center>![spectrum][loew]</center>
Source: L. M. Loew [Design and Use of Organic Voltage Sensitive Dyes](http://www.newbooks-services.de/MediaFiles/Texts/8/9781441965578_Excerpt_001.pdf) (ungated pdf).<br/><br/>

There is a slight skew in the above curves, as would be expected if the frequency spectrum, i.e., relative fluorescence as a function of frequency rather than as a function of wavelength, were unimodal and symmetric. The shapes in the figure can, in fact, be approximated by using a Gaussian frequency spectrum with a mean corresponding to an (arbitrary) peak wavelength of 675 nm, i.e., a frequency of $\frac{300}{675}\:PHz$, and supposing that 150 nm corresponds to 2.5 standard deviations in frequency, i.e.,$$sd = \frac{1}{2.5}\left(\frac{300}{675-125}-\frac{300}{675}\right).$$

```{r fig.align='center', fig.show='hold', fig.width=7.5, echo=FALSE}
fmean <- 300/675
fsd <- (300/525 - fmean)/2.5
# Changing the variable from frequency to wavelength introduces the factor of 300/lambda^2
spec <- function(lambda, shift=0)dnorm(300/(lambda-shift), fmean, fsd)*300/(lambda-shift)^2
lambda <- seq(550, 850, length.out = 100)
plot(lambda, spec(lambda+20), type='l', col="gray", lwd=3, xlab="wavelength (nm)", ylab="", yaxt="n", bty='n', main = "Stylized VSD spectra")
lines(lambda, spec(lambda), col="black", lwd=3)
legend('topright', c("Polarized", "Depolarized"), lwd=3, col=c("black", "gray"), bty='n')
```

Peak emission and absorption wavelengths for various lipid-bound chromophores are given in Leslie Loew's review, (*ibid*.) Spectrum bandwidths are not indicated in the review but widths of several hundred nanometers seem common elsewhere in the literature. Hence, my choice of +/- 150 nm above.

A convenient model based on the above can be created with the following code (which also appears in R/utilities.R.) Its use and interpretation are discussed below. 

```{r}
# peak -- peak wavelength in nanometers
# width -- width (support) of the spectrum in nanometers
# Returns a function which gives relative fluorescence or absorption as a function of
# wavelength and shift due to depolarization.
vsdSpectrum <- function(peak, width=300){
   # Create a function for a curve with a 675 nm peak
  # NOTE: the factor 300/lambda^2 comes from a change of variable
  # from frequency to wavelength.
  fsd <- (300/(675-.5*width) - 300/675)/2.5
  fcal <- function(lambda)dnorm(300/lambda, 300/675, fsd)*300/lambda^2
  # Return a function which translates the calibrated curve
  # to the correct peak wavelength and shift, and normalizes
  # it so its peak value is 1.
  function(lambda, depolarization_shift=0){
    fcal(lambda + depolarization_shift + 675 - peak)/fcal(675)
  }
}
```

### Interpretation

For simulations, I'll assume that voltage-induced variations in an absorption spectrum correspond to proportional variations in a dye's absorption coefficient, $\mu_a.$ Thus, if depolarization causes a -15% change in the absorption spectrum at a certain wavelength, I'll assume the absorption coefficient changes by -15% as well.

I've [estimated](http://brain-initiative.github.io/know_brainR/Rmds/absorption.html) that absorption coefficients of 1 micromolar concentrations of VSD will be about the same as those of gray matter. In this case, a variation of -15% in a dye's coefficient would result in a -7.5% variation in the region's.

After dye absorbs a photon, it will lose some energy through vibration then either emit a photon at a longer wavelength or relax entirely through heat. If it re-emits a photon, it will do so at a longer wavelength as governed by its *emission* spectrum. Interpreting its emission spectrum as a curve of relative emission rates, the mean wavelength of re-emission can be calculated as if the spectrum were a probability distribution:$$E\left(\lambda_{em}\right) = \frac{\int_{\lambda_{abs}}^{\infty}\lambda S_{em}(\lambda)d\lambda}{\int_{\lambda_{abs}}^{\infty}S_{em}(\lambda)d\lambda}$$

One of the chromophores listed in Leslie Loew's review (*ibid*.) has a lipid-bound absorption spectrum with peak at 678 nm, and a lipid-bound emission spectrum with peak at 770 nm. If, in a polarized state, the dye is excited by an external source at 750 nm, the average emission wavelength will be 828 nm as shown in the figure.

```{r fig.align='center', fig.show = 'hold', echo=FALSE, fig.width=8}
S_abs <- vsdSpectrum(678)
S_em <- vsdSpectrum(770)
lambda <- seq(500, 1000, by=10)
idx <- lambda > 750
n <- sum(idx)
matplot(lambda, cbind(S_abs(lambda), S_em(lambda)), type='n', xlab="wavelength (nm)", ylab = "", yaxt='n', lty=1, lwd = 3, col=c("blue", "red"), main="Styryl dye chromophore [7–1–4]\nPolarized")
polygon(c(lambda[idx], lambda[idx][n:1], 750, 750), c(rep(0, n), S_em(lambda[idx][n:1]), S_em(750), 0), col="pink", border=NA)
lines(lambda, S_em(lambda), lwd=3, col="red")
lines(lambda, S_abs(lambda), lwd=3, col="blue")
segments(750, 0, 750, S_abs(750), lty=2, lwd=7, col="blue")
avg_em <- integrate(function(lambda)lambda*S_em(lambda), 750, 1500)$value/integrate(S_em, 750, 1500)$value
segments(avg_em, 0, avg_em, S_em(avg_em), lty=2, lwd=7, col="red")
legend('topright', c(expression(S[abs]), expression(S[em]), "excitation (750 nm)", "avg emission (828 em)"), col=c("blue", "red"), lwd=c(3, 3, 7, 7), lty=c(1, 1, 2, 2))
```

Since energy is inversely proportional to wavelength re-emission will involve an energy loss of $1-750/878 \approx 0.094$, a little less than 10%. A dye with a quantum yield of 95%, will fail to re-emit 5% of absorbed photons, which would total a little less than 15% loss to vibration and heat. In a photoacoustic scenario, since the speed of thermal diffusion in tissue should be much less than the speed of sound, I believe most of this ~15% would dissipate sonically.

If depolarization shifts both spectra to the left by 30 nm, the excitation remains the same, but the average re-emission wavelength decreases to 820 nm, resulting in re-emission loss of $1 - 750/820 \approx 0.085$, about 1% less than before. (Of course, absorption will decrease, in this case, by about 32%, so total emitted energy, optical plus acoustic, will decrease by a similar amount.)

```{r fig.align='center', fig.show = 'hold', echo=FALSE, fig.width=8}
shft <- 30
matplot(lambda, cbind(S_abs(lambda, shft), S_em(lambda, shft)), type='n', xlab="wavelength (nm)", ylab = "", yaxt='n', lty=1, lwd = 3, col=c("blue", "red"), main="Styryl dye chromophore [7–1–4]\nDepolarized")
polygon(c(lambda[idx], lambda[idx][n:1], 750, 750), c(rep(0, n), S_em(lambda[idx][n:1], shft), S_em(750,shft), 0), col="pink", border=NA)
lines(lambda, S_em(lambda, shft), lwd=3, col="red")
lines(lambda, S_abs(lambda, shft), lwd=3, col="blue")
segments(750, 0, 750, S_abs(750, shft), lty=2, lwd=7, col="blue")
text(750, S_abs(750, shft), "-32%", col="blue", pos=2, cex=1.5)
avg_em <- integrate(function(lambda)lambda*S_em(lambda,shft), 750, 1500)$value/integrate(function(lambda)S_em(lambda, shft), 750, 1500)$value
segments(avg_em, 0, avg_em, S_em(avg_em, shft), lty=2, lwd=7, col="red")
legend('topright', c(expression(S[abs]), expression(S[em]), "excitation (750 nm)", "avg emission (820 nm)"), col=c("blue", "red"), lwd=c(3, 3, 7, 7), lty=c(1, 1, 2, 2))
```

 
[loew]: ../images/loew.png