---
title: "Spatial Resolution"
author: "W. Bauer"
date: "12/13/2014"
output: html_document
---

Reference: Natalie Baddour, ["Operational and convolution properties of two-dimensional Fourier transforms in polar coordinates"](http://www.researchgate.net/publication/26711481_Operational_and_convolution_properties_of_two-dimensional_Fourier_transforms_in_polar_coordinates/file/e0b4952a2409f508bc.pdf)(pdf), J. Opt. Soc. Am. A, Vol. 26, No. 8, August 2009.

For an extremely simplified model, [Estimated Loss and Dispersion of Photons](loss.html) derived a formula for the fraction photons emitted at a depth of $b$, which will reach a 1 mm square area of the "scalp" at a distance $r$ from a point above the point of emission. For an infinitesimal area, which is $rdrd\theta$ in polar coordinates, the corresponding formula is$$f_b(r)rdrd\theta = \frac{b}{4 \pi \left(b^2 + r^2\right)^{\frac{3}{2}}} 0.92^{\sqrt{b^2 + r^2}}rdrd\theta.$$ 

```{r echo=FALSE, fig.show='hold', fig.align='center'}
# Estimates fraction of photons which will reach the scalp at a point r mm distant
# from the point directly over their origin, assuming their origin is b mm beneath
# the scalp.
f_b <- function(r, b){
  r1 <- sqrt(r^2+b^2)
  b*.92^r1/(4*pi*r1^1.5)
}
```

The same formula holds for the fraction of photons reaching a point on the "scalp" from an infinitesimal area of the "cerebral cortex" at a lateral distance, $r,$ and depth, $b.$ Thus the number of photons impinging on a point on the "scalp" is given by centering $f_b(r)drd\theta$ on that point and integrating its product with the pattern of emissions $b$ mm directly beneath. Repeating this process for every point on the "scalp" amounts to a two dimensional convolution of the product of $f_b(r)$ with the pattern of emissions at the "cerebral cortex." In other words, this simplified model is a standard linear system, and $f_b(r)$ specifies a spatial low-pass filter. Its Fourier transform, which diagonalizes the filter, hence gives its eigenvalues, is therefore of interest. The magnitude of its eigenvalues relative to measurement error and other noise levels determine the accuracy with which it can be inverted.

The 2D Fourier transform of a real valued, radially symmetric function is also a real valued radially symmetric function. Letting $\rho$ represent spatial frequency in cycles per mm, the unitary (i.e., square norm or power preserving) version of the 2D Fourier transform is,$$\hat{f}(2 \pi \rho) = \int_0^{\infty}f_b(r)J_0(2 \pi \rho r)rdr$$where $J_0$ is the zero^th^ order Bessel function the first kind.

```{r echo=FALSE, fig.show='hold', fig.align='center'}
# Computes the Fourier transform of f_b at spatial frequency freq. Both the above
# and its Fourier transform are radially symmetric.
xform <- function(freq, b){
  integrand <- function(r){r*f_b(r,b)*besselJ(2*pi*freq*r,0)}
  integrate(integrand, 0, Inf)
}
```

Below is a table of spatial frequencies, $\rho,$ and corresponing eigenvalues, equivalently Fourier transform values, estimated by numerical integration of this formula. Note that for spatial frequencies above .1 cycles/mm, estimated numerical integration errors are as large as the eigenvalues themselves. Thus, the most that can be said using this approach is that eigenvalues corresponding to .2 cycles/mm and above are of magnitude $10^{-5}$ or smaller.

Since spatial resolution corresponds roughly to the inverse of spatial frequency, these calculations suggest adequate spatial resolutions of around 1/.1 mm, or 1 cm.

Spatial Frequency | Eigenvalue | Estimated Error
------------------ | ---------- | -----
0 | 0.9601971 | < 6e-05
.1 | 0.001150474 | < 6.3e-05
.2 | 6.233459e-06 | < 1e-04
.3 | 1.167508e-05 | < 7.7e-05
.33 | 1.215592e-05 | < 6.3e-05
