---
title: "Voxel Level Statistics"
author: "gina"
date: "05/14/2015"
output: html_document
---

Since heterogeneity of tissue properties is essential to image recovery, anatomical or physiological, it is important to understand its effects both on optical and acoustic propagation. Here we consider its effects on optical propagation at a cubic mm voxel level. In particular, we estimate the probabilites of scattering, absorption, and hitting a voxel boundary as they depend on a voxel's scattering and absorption coefficients. We assume that photons are distributed uniformly in the voxel volume and are traveling in random directions.

For the BrainWeb phantom's eleven tissue types, scattering is significantly more probable than either of the other two mutually exclusive possibilities, absorption or hitting a boundary. This has complicated previous attempts to estimate the probabilities, since relatively few absorption events or boundary encounters occur. Here we remedy the problem by first getting reliable probabilities for scattering, then getting probabilities for the other two events *conditional* on not scattering.

The general strategy is to generate three distances in each random trial: distance to a scattering event, distance to an absorption event, and distance to a boundary in the direction of travel. Scattering and absorption distances are distributed exponentially with parameters, $\mu_s$ and $\mu_a$, respectively, so generating these at random is standard fare. Generating distance to a boundary at random requires, first, to generate a random position in a cube, next, to generate a random direction (i.e., a random point on the unit sphere,) and finally, to calculate the distance in that direction to the first boundary encountered. The smallest of the three distances determines the which event occurs. The relevant source code is in [alpha.R](../R/alpha.R) and [utilities.R](../R/utilities.R).

```{r}
source("../R/alpha.R")
source("../R/utilities.R")
```

### Scatter

We first estimate scattering probabilties for the 11 tissue types. Since the scattering, as opposed to being absorbed or hitting a boundary, is a Bernoulli trial, it is relatively easy to estimate standard errors. Based on standard errors, we use $10^6$ trials.

```{r eval=FALSE}
tissue_properties <- read.csv("../data/tissue_properties.csv")
lg <- c("Background", "CSF", "Gray Matter", "White Matter", "Fat", "Muscle", "Muscle/Skin", "Skull", "Vessels", "Around fat", "Dura mater", "Bone marrow")
ranseed <- 1431632596 # system time when code was written
scatter <- data.frame(id=1:11, tissue=lg[2:12], mu_a=tissue_properties[2:12, "mu_a"],
                      mu_s = tissue_properties[2:12, "mu_s"], p_scatter=NA, std_err_scatter=NA)
for(i in 1:11){
  mua <- scatter[i, "mu_a"]
  mus <- scatter[i, "mu_s"]
  temp <- pscatter(mua,mus,1e+6,ranseed+i)
  scatter[i, "p_scatter"] <- temp$prscatt
  scatter[i, "std_err_scatter"] <- temp$se_scatt
}

scatter
```

```{r echo=FALSE}
scatter <- read.table("../data/pscatter.csv", sep=",", header=TRUE)
scatter
```

### Absorption and boundary encounters

With reliable estimates of probabilities of (not) scattering in hand, we can compute probabilities of absorption and hitting boundaries, conditional on not scattering.

```{r eval=FALSE} 
ranseed <- 1431635764
other <- data.frame(id=1:11, tissue=lg[2:12], mu_a=tissue_properties[2:12, "mu_a"],
                      mu_s = tissue_properties[2:12, "mu_s"], cp_absorb=NA, cp_bound=NA, std_err=NA)
for(i in 1:11){
  mua <- scatter[i, "mu_a"]
  mus <- scatter[i, "mu_s"]
  temp <- pother(mua,mus,1e+6,ranseed+i)
  other[i, "cp_absorb"] <- temp$cpabsorb
  other[i, "cp_bound"] <- temp$cpbound
  other[i,"std_err"] <- temp$se
}

other
```

```{r echo=FALSE}
other <- read.table("../data/conditional_pabs_pbdry.csv", sep=",", header=TRUE)
other
```

To get unconditional probabilities from the conditional probabilities just estimated, we must multiply each by the probability of not scattering. Making a table of final results:

```{r eval=FALSE}
vox_probs <- scatter[,1:5]
vox_probs[,"p_absorb"] <- (1-vox_probs[,"p_scatter"])*other[,"cp_absorb"]
vox_probs[,"p_boundary"] <- (1-vox_probs[,"p_scatter"])*other[,"cp_bound"]
vox_probs
```

```{r echo=FALSE}
vox_probs <- read.table("../data/vox_probs.csv", sep=",", header=TRUE)
vox_probs
```

### Crossing boundaries

To simulate diffusion of photons across voxel boundaries we need, in addition to the probabilities of hitting one of the 6 voxel boundaries, the probabilities of crossing a boundary once it is encountered. This depends on the indices of refraction on either side of the particular (1 of 6 at random) voxel boundary encountered. Function `prob_refl` in [utilities2.R](../R/utilities2.R) calculates the probability of reflection based on the angle of incidence and the two indices. Since in our case angles of incidence are regarded as random, we integrate over all angles of incidence (0 to $\pi/2$) regarding each as equally likely, hence dividing the integral by $\pi/2$. Since we are interested in the probability of *not* reflecting, we subtract the result from 1.

```{r eval=FALSE}
# A wrapper for prob_refl to make it friendly to R's integrate function
prefl <- function(theta, n1, n2){
  n1 <- rep(n1, length(theta))
  n2 <- rep(n2, length(theta))
  prob_refl(theta, n1, n2)
}

# Create a data frame to hold answers
bdry_probs <- data.frame(matrix(0,11,13))
names(bdry_probs) <- c("source_tissue", lg)
bdry_probs[,1] <- lg[2:12]
# Compute probabilities for each source/target tissue pair
for(i in 1:11){
  for(j in 1:12){
    n1 <- tissue_properties[i+1,"n"]
    n2 <- tissue_properties[j, "n"]
    bdry_probs[i, j+1] <- 1-(2/pi)*integrate(prefl, 0, pi/2, n1, n2)$value
  }
}
bdry_probs[,1:5]
bdry_probs[,c(1,6:9)]
bdry_probs[,c(1,10:12)]
```

```{r echo=FALSE}
bdry_probs <- read.table("../data/boundary_crossing_probs.csv", sep=",", header=TRUE)
bdry_probs[,1:5]
bdry_probs[,c(1,6:9)]
bdry_probs[,c(1,10:12)]
```

### Conclusion

We now have estimates for the proportion of a voxel's photons which will be absorbed per "step", the proportion which will encounter a boundary (1 of 6,) and the proportion of those encountering a boundary which will cross it. A diffusive model based on such proportions will a) obey conservation of energy and b) model effects of tissue heterogeneity.

A "step" refers to a distance between successive events on a photon's path, an "event" being scattering, absorption, or a boundary encounter. Based on average speed of light in tissue, there are roughly 500 steps per nanosecond.
