---
title: "Voxel Level Statistics"
author: "gina"
date: "05/14/2015"
output: html_document
---

Since heterogeneity of tissue properties is essential to image recovery, anatomical or physiological, it is important to understand its effects both on optical and acoustic propagation. Here we consider its effects on optical propagation at a cubic mm voxel level. In particular, we estimate the probabilites of scattering, absorption, and hitting a voxel boundary as they depend on a voxel's scattering and absorption coefficients. We assume that photons are distributed uniformly in the voxel volume and are traveling in random directions.

For the BrainWeb phantom's eleven tissue types, scattering is significantly more probable than either of the other two mutually exclusive possibilities, absorption or hitting a boundary. This has complicated previous attempts to estimate the probabilities, since relatively few absorption events or boundary encounters occur. Here we remedy the problem by first getting reliable probabilities for scattering, then getting probabilities for the other two events *conditional* on not scattering.

The general strategy is to generate three distances in each random trial: distance to a scattering event, distance to an absorption event, and distance to a boundary in the direction of travel. Scattering and absorption distances are distributed exponentially with parameters, $\mu_s$ and $\mu_a$, respectively, so generating these at random is standard. Generating distance to a boundary at random required first to generate a random position in a cube, the to generate a random direction (i.e., a random point on the unit sphere,) and finally to calculate the distance in that direction to the first boundary encountered. The smallest of the three distances randomly generated determines the type of event which actually occurred. The relevant source code is in [alpha.R](../R/alpha.R) and [utilities.R](../R/utilities.R).

```{r}
source("../R/alpha.R")
source("../R/utilities.R")
```

We first estimate scattering probabilties for the 11 tissue types. Since the scattering, as opposed to being absorbed or hitting a boundary, is a Bernoulli trial, it is relatively easy to estimate standard errors. Based on standard errors, we use $10^6$ trials.

```{r eval=FALSE}
tissue_properties <- read.csv("../data/tissue_properties.csv")
lg <- c("Background", "CSF", "Gray Matter", "White Matter", "Fat", "Muscle", "Muscle/Skin", "Skull", "Vessels", "Around fat", "Dura mater", "Bone marrow")
ranseed <- 1431632596 # system time when code was written
scatter <- data.frame(id=1:11, tissue=lg[2:12], mu_a=tissue_properties[2:12, "mu_a"],
                      mu_s = tissue_properties[2:12, "mu_s"], p_scatter=NA, std_err_scatter=NA)
for(i in 1:11){
  mua <- scatter[i, "mu_a"]
  mus <- scatter[i, "mu_s"]
  temp <- pscatter(mua,mus,1e+6,ranseed+i)
  scatter[i, "p_scatter"] <- temp$prscatt
  scatter[i, "std_err_scatter"] <- temp$se_scatt
}
```

Given reliable estimates of probabilities of (not) scattering, we can now compute probabilities of absorption and hitting boundaries, conditional on not scattering.

```{r eval=FALSE} 
ranseed <- 1431635764
other <- data.frame(id=1:11, tissue=lg[2:12], mu_a=tissue_properties[2:12, "mu_a"],
                      mu_s = tissue_properties[2:12, "mu_s"], cp_absorb=NA, cp_bound=NA, std_err=NA)
for(i in 1:11){
  mua <- scatter[i, "mu_a"]
  mus <- scatter[i, "mu_s"]
  temp <- pother(mua,mus,1e+6,ranseed+i)
  other[i, "cp_absorb"] <- temp$cpabsorb
  other[i, "cp_bound"] <- temp$cpbound
  other[i,"std_err"] <- temp$se
}
```

To get unconditional probabilities from the conditional probabilities just estimated, we must multiply by the probability of not scattering. Making a table of final results:

```{r eval=FALSE}
vox_probs <- scatter[,1:5]
vox_probs[,"p_absorb"] <- (1-vox_probs[,"p_scatter"])*other[,"cp_absorb"]
vox_probs[,"p_boundary"] <- (1-vox_probs[,"p_scatter"])*other[,"cp_bound"]
```

